The GitHub Open Source Development Process \author{Kevin Peterson\\ Department of Information Technology\\ Clinical Informatics Support Systems\\ Mayo Clinic\\ 200 1st Street\\ Rochester, MN 55905\\ \small \texttt{peterson.kevin@mayo.edu} }

### 2013-04-28 paper/main.tex
Finished another round of typo checks.

GitHub\footnote{https://github.com/} is a popular code hosting website that uses Git Source Code Management (SCM).\footnote{http://git-scm.com/} Along with hosting software repositories, GitHub incorporates the social aspects of development. GitHub users have community-visible profiles, and user actions can be tracked and followed by other community actions. This type of social integration of a user's identity and actions is unique to GitHub~\cite{dabbish2012social}.

@@

This hypothesis is, ultimately, an examination of ``Linus Law"~\cite{raymond1999cathedral} in the context of GitHub. In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.\\

@@

\noindent \textbf{Research Question 2: How is GitHub changing the OSS process?}\\
The social aspects of GitHub are an important part of the development experience~\cite{dabbish2012social}. User interactions and social pressures can drive OSS development in interesting ways. Social aspects of OSS development have existed before GitHub, with mailing lists~\cite{mockus2000case}, gift giving mechanisms~\cite{bergquist2008power}, and other collaboration tools. What is GitHub doing to facilitate change in OSS development, and how does it see OSS development changing moving forward?\\

\noindent \textbf{Research Question 3: Can GitHub be used for more than code artifacts?}\\
The Object Management Group\textsuperscript{\textregistered}\footnote{http://www.omg.org} is a computer industry consortium focused on technology standards. Potential standard submission teams must go through a vetting process~\cite{kobryn1999uml} which allows industry representatives to provide feedback on the standard. Gathering and recording this feedback is a challenge, but GitHub's issue tracking system may be able to streamline the process. This just one example of how GitHub is being used for non-code artifacts. Are organizations exploring this type of work flow?

@@

To gather the metrics, the sets of all individual \textit{Attribute} across all \textit{Repositories} are processed by some aggregate function $f$. The input of this function is the union of all sets of a particular \textit{Attribute} over each \textit{Repository} in the sample. For this analysis, the aggregate functions are: \textsc{mean}, \textsc{min}, \textsc{max}, \textsc{stddev}. This analysis is applied to all repositories in the sample $S$ for which data was collected. Each attribute $\alpha$ is analyzed independently, so given:
\[ g(\alpha) =f \left( \bigcup\limits_{r = 0}^{|S|}  A_{r,\alpha} \right) \]
each individual attribute is processed via $g(\alpha)$ and will result in the output of the given aggregate function. The result of each aggregate function when applied to the entire sample set of \textit{Repositories} is shown in Table~\ref{table:summary_stats}.

@@

The average is calculated using the following method. Let $d(x)$ be a function that converts seconds to a Date, and let $s(x)$ be a function that converts a Date to seconds. Let $I$ equal the Set of \textit{Issues} in a given repository, with elements comprised of tuples $(\tau, \tau^\prime)$, where $\tau$ is \textit{Issue Open Date} and $\tau^\prime$ is \textit{Issue Close Date}. We calculate \textit{Issue Close Time} by:

@@

\noindent \textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
No significant correlation was found between the number of repository forks and issue resolution time. It is important to note, however, that issue resolution time is not necessarily related to software product quality in the context of GitHub. We can assert that for several reasons. \textit{First}, GitHub issues are not necessarily product defects. Issues may be feature requests, comments, or general support questions. \textit{Second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all. \textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub. This, among other things, suggests that GitHub may be facilitating testing by making it more \emph{triable} by lowering entry barriers, and makng the process as a whole more \emph{observable}. This aspect is something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric~\cite{phamcreating}.

@@

\begin{figure}
\includegraphics[height=3in,width=3in]{images/committers_percentage_pie_chart.png}
\caption{Distribution of individual committer commit percentage of a repository}
\label{fig:committers_percentage_pie_chart}
\end{figure}

@@

According to GitHub Staff member John Britton, lowering the barrier to participation is an important cultural shift:
\begin{quote}
GitHub moved the open source community away from a permission culture. Prior to GitHub, contributing to an open source project was very difficult and often involved asking permission. By allowing people to fork a repository and make changes without asking for permission, GitHub has made contributing to someone else's project much easier than it has ever been. By submitting a Pull Request with suggested changes, project maintainers and contributors have a place to discuss changes to code that is already written and to refine it.~\cite{github_interview}
\end{quote}

@@

\noindent \textbf{Research Question 3: Can GitHub be used for more than code artifacts?}\\
The OMG\textsuperscript{\textregistered} standardization process is a process by which products strive to produce an interoperability standard by way of an industry standards consortium~\cite{omgTechProcess}.
The standardization process itself aims to, at its conclusion, produce a set of model artifacts containing sufficient detail such that when interpreted and implemented, produce interoperable software.
At a conceptual level, this can be thought of as a software requirements engineering process as much as a standardization process. If so, of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,~\cite{sommerville1998requirements} we can show that GitHub is an appropriate forum for many of them.

@@

The data does not, however, reflect some other traditional aspects of OSS development. Notably, it was observed that data gathered for Hypothesis 2 and Hypothesis 3 seemed contrary to what the Apache Server report concluded~\cite{mockus2000case}. A possible explanation is the amount of variability in the data collected. Summary Data Assertion~\ref{eq:variance} (Section~\ref{sec:summary_statistics}), which is derived from the summary data shown in Table~\ref{table:summary_stats}, suggests large amounts of variety in the observed variables. This may suggest a wide range of repository purposes, or specifically, that not all repositories on GitHub are intended to be consumable software artifacts for users, or ``customers.'' A large number of repositories may not be intended for customer consumption, but merely as example code, experimentation, or exercises. Given this, and specific reasons outlined in the hypotheses themselves, we cannot state any definitive conclusions regarding these two hypotheses. More research into categorizing GitHub repositories is needed.

>>>>

GitHub\footnote{https://github.com/} is a popular code hosting website that uses Git Source Code Management (SCM).\footnote{http://git-scm.com/} Along with hosting software repositories, GitHub incorporates the social aspects of development. GitHub users have community-visible profiles, and user actions can be tracked and followed by other community members. This type of social integration of a user's identity and actions is unique to GitHub~\cite{dabbish2012social}.

@@

This hypothesis is, ultimately, an examination of ``Linus Law"~\cite{raymond1999cathedral} in the context of GitHub. In other words, the more exposure the code gets (in terms of \emph{forks}), the easier bugs will be to find and fix.\\

@@

\noindent \textbf{Research Question 2: How is GitHub changing the OSS process?}\\
The social aspects of GitHub are an important part of the development experience~\cite{dabbish2012social}. User interactions and social pressures can drive OSS development in interesting ways. Social aspects of OSS development have existed, however, before GitHub, with mailing lists~\cite{mockus2000case}, gift giving mechanisms~\cite{bergquist2008power}, and other collaboration tools. What is GitHub doing to facilitate and capitalize on the social aspects of OSS development, and how might this influence the OSS development process?\\

\noindent \textbf{Research Question 3: Can GitHub be used for more than code artifacts?}\\
The Object Management Group\textsuperscript{\textregistered}\footnote{http://www.omg.org} is a computer industry consortium focused on technology standards. Potential standard submission teams must go through a vetting process~\cite{kobryn1999uml} which allows industry representatives to provide feedback on the standard. Gathering and recording this feedback is a challenge, but GitHub's issue tracking system may be able to streamline the process. This is just one example of how GitHub is being used for non-code artifacts. Are organizations exploring this type of work flow?

@@

To analyze the metrics, the sets of all individual \textit{Attribute} across all \textit{Repositories} are processed by some aggregate function $f$. The input of this function is the union of all sets of a particular \textit{Attribute} over each \textit{Repository} in the sample. For this analysis, the aggregate functions are: \textsc{mean}, \textsc{min}, \textsc{max}, \textsc{stddev}. This analysis is applied to all repositories in the sample $S$ for which data was collected. Each attribute $\alpha$ is analyzed independently, so given:
\[ g(\alpha) =f \left( \bigcup\limits_{r = 0}^{|S|}  A_{r,\alpha} \right) \]
each individual attribute is processed via $g(\alpha)$ and will result in the output of the given aggregate function. The result of each aggregate function when applied to the entire sample set of \textit{Repositories} is shown in Table~\ref{table:summary_stats}.

@@

The average \textit{issue close time} is calculated using the following method: Let $d(x)$ be a function that converts seconds to a Date, and let $s(x)$ be a function that converts a Date to seconds. Let $I$ equal the Set of \textit{Issues} in a given repository, with elements comprised of tuples $(\tau, \tau^\prime)$, where $\tau$ is \textit{Issue Open Date} and $\tau^\prime$ is \textit{Issue Close Date}. We calculate \textit{Issue Close Time} by:

@@

\noindent \textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
No significant correlation was found between the number of repository forks and issue resolution time. It is important to note, however, that issue resolution time is not necessarily related to software product quality in the context of GitHub. We can assert that for several reasons. \textit{First}, GitHub issues are not necessarily product defects. Issues may be feature requests, comments, or general support questions. \textit{Second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all. \textit{Third}, Cultural shifts in how testing is approached may have significant impact on quality. Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub, which suggests that GitHub may be facilitating testing by making it more \emph{triable} by lowering entry barriers, and makng the process as a whole more \emph{observable}. This aspect is something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric~\cite{phamcreating}.

@@

\begin{figure}
\includegraphics[height=3in,width=3in]{images/committers_percentage_pie_chart.png}
\caption{Distribution of individual committer contribution percentage to a repository}
\label{fig:committers_percentage_pie_chart}
\end{figure}

@@

According to GitHub Education Liaison John Britton, lowering the barrier to participation is an important cultural shift:
\begin{quote}
GitHub moved the open source community away from a permission culture. Prior to GitHub, contributing to an open source project was very difficult and often involved asking permission. By allowing people to fork a repository and make changes without asking for permission, GitHub has made contributing to someone else's project much easier than it has ever been. By submitting a Pull Request with suggested changes, project maintainers and contributors have a place to discuss changes to code that is already written and to refine it.~\cite{github_interview}
\end{quote}

@@

\noindent \textbf{Research Question 3: Can GitHub be used for more than code artifacts?}\\
The OMG\textsuperscript{\textregistered} standardization process is a process by which products strive to produce an interoperability standard by way of an industry standards consortium~\cite{omgTechProcess}.
The standardization process itself aims to, at its conclusion, produce a set of model artifacts containing sufficient detail such that when interpreted and implemented, produce interoperable software.
At a conceptual level, this can be thought of as a software requirements engineering process as much as a standardization process. If so, of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management~\cite{sommerville1998requirements}, we can show that GitHub is an appropriate forum for many of them.

@@

The data does not, however, reflect some other traditional aspects of OSS development. Notably, it was observed that data gathered for Hypothesis~2 (\textit{As the number Repository Forks increase, the Issue Resolution Time will decrease}) and Hypothesis~3 (\textit{There will be more Issue Reporters than Committers by an order of magnitude}) seemed contrary to what the Apache Server report concluded~\cite{mockus2000case}. A possible explanation is the amount of variability in the data collected. Summary Data Assertion~\ref{eq:variance} (Section~\ref{sec:summary_statistics}), which is derived from the summary data shown in Table~\ref{table:summary_stats}, suggests large amounts of variety in the observed variables. This may suggest a wide range of repository purposes, or specifically, that not all repositories on GitHub are intended to be consumable software artifacts for users, or ``customers.'' A large number of repositories may not be intended for customer consumption, but merely as example code, experimentation, or exercises. Given this, and specific reasons outlined in the hypotheses themselves, we cannot state any definitive conclusions regarding these two hypotheses. More research into categorizing GitHub repositories is needed.
